---
  title: "GHG Fieldsheets quality check"
  author: "Camille Minaudo"
  date: "2024-05-24"
  output:
  html_document: default
  pdf_document: default
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load packages and source functions

```{r loading packages and functions, echo=F, message=FALSE, warning=FALSE}
rm(list = ls()) # clear workspace
cat("/014") # clear console


# ---- packages ----
library(tidyverse)
library(readxl)
library(lubridate)
library(zoo)
library(ggplot2)
library(grid)
library(egg)
require(dplyr)
require(purrr)
require(sf)


source(paste0(dirname(dirname(rstudioapi::getSourceEditorContext()$path)),"/functions/get_unix_times.R"))
source(paste0(dirname(dirname(rstudioapi::getSourceEditorContext()$path)),"/functions/read_GHG_fieldsheets.R"))

# set directories
dropbox_root <- "C:/Users/Camille Minaudo/Dropbox/RESTORE4Cs - Fieldwork/Data"
fieldsheetpath <- paste0(dropbox_root,"/GHG/Fieldsheets")

```

## Load all fieldsheets in the Dropbox or a chosen sampling

```{r Load all fieldsheets in the Dropbox, echo=F, message=FALSE, warning=FALSE}

# ---- List GHG chamber fieldsheets in Dropbox and read them ---
# list filenames
myfieldsheets_list <- list.files(fieldsheetpath, pattern = "Fieldsheet-GHG.xlsx", all.files = T, full.names = T, recursive = T)

# uncomment the following lines if you want to check fieldsheets only for a given sampling
# sampling <- "S3"
# i <- grep(pattern = sampling, x = myfieldsheets_list) # selecting the files corresponding to the selected sampling campaign
# myfieldsheets_list <- myfieldsheets_list[i]

# Read all fieldsheets and put them in a single dataframe
fieldsheet <- read_GHG_fieldsheets(myfieldsheets_list)
fieldsheet_flagged <- fieldsheet[0,]
fieldsheet_flagged$flag <- ""

append2table <- function(fieldsheet, fieldsheet_flagged, ind, flag_txt){
  fieldsheet_sel <- fieldsheet[ind,]
  fieldsheet_sel$flag <- flag_txt
  fieldsheet_flagged <- rbind(fieldsheet_flagged, fieldsheet_sel)
  
  return(fieldsheet_flagged)
}


message("loaded!")
```


## Quality check 1 - Start/Stop times

```{r Start/Stop times, echo=F}
ind_erronous_times <- which(fieldsheet$unix_stop < fieldsheet$unix_start)

if(length(ind_erronous_times)>0){
  message("the following rows show erronous start/stop times")
  as.data.frame(fieldsheet[ind_erronous_times,c("pilot_site","subsite","person_sampling","plot_id","start_time","end_time")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_erronous_times, "erronous start/stop time")
}

# rows with duration suspicious
ind_duration_neg <- which(fieldsheet$unix_stop-fieldsheet$unix_start < 0)
ind_duration_tooshort <- which(fieldsheet$unix_stop-fieldsheet$unix_start < 3*60)
ind_duration_toolong <- which(fieldsheet$unix_stop-fieldsheet$unix_start > 15*60)

ind_duration <- c(ind_duration_neg, ind_duration_tooshort, ind_duration_toolong)

fieldsheet$duration <- fieldsheet$unix_stop-fieldsheet$unix_start

if(length(ind_duration)>0){
  message("the following rows show suspicious incubation duration")
  as.data.frame(fieldsheet[ind_duration,c("pilot_site","subsite","person_sampling","plot_id","start_time","end_time","duration")])
  
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_duration, "suspicious duration")
}
```


## Quality check 2 - suspicious CO2 values

```{r CO2 values, echo=F}

ind_suspicious_co2_initial <- which(fieldsheet$initial_co2 > 1000)
ind_suspicious_co2_final <- which(fieldsheet$final_co2 > 1000)
ind_suspicious_co2 <- unique(c(ind_suspicious_co2_initial, ind_suspicious_co2_final))

if(length(ind_suspicious_co2)>0){
  message("the following rows show suspiciously HIGH CO2 levels")
  as.data.frame(fieldsheet[ind_suspicious_co2,c("pilot_site","subsite","person_sampling","plot_id","initial_co2","final_co2")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_suspicious_co2, "suspiciously high CO2")
}

ind_suspicious_co2_initial <- which(fieldsheet$initial_co2 < 200)
ind_suspicious_co2_final <- which(fieldsheet$final_co2 < 200)
ind_suspicious_co2 <- unique(c(ind_suspicious_co2_initial, ind_suspicious_co2_final))


if(length(ind_suspicious_co2)>0){
  message("the following rows show suspiciously LOW CO2 levels")
  as.data.frame(fieldsheet[ind_suspicious_co2,c("pilot_site","subsite","person_sampling","plot_id","initial_co2","final_co2")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_suspicious_co2, "suspiciously low CO2")
}
```



## Quality check 3 - suspicious CH4 values

```{r CH4 values, echo=F}
ind_suspicious_ch4_initial <- which(fieldsheet$initial_ch4 > 2000)
ind_suspicious_ch4_final <- which(fieldsheet$final_ch4 > 2000)
ind_suspicious_ch4 <- unique(c(ind_suspicious_ch4_initial, ind_suspicious_ch4_final))

if(length(ind_suspicious_ch4)>0){
  message("the following rows show suspiciously HIGH methane levels")
  as.data.frame(fieldsheet[ind_suspicious_ch4,c("pilot_site","subsite","person_sampling","plot_id","initial_ch4","final_ch4")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_suspicious_ch4, "suspiciously high CH4")
}

ind_suspicious_ch4_initial <- which(fieldsheet$initial_ch4 < 1)
ind_suspicious_ch4_final <- which(fieldsheet$final_ch4 < 1)
ind_suspicious_ch4 <- unique(c(ind_suspicious_ch4_initial, ind_suspicious_ch4_final))

if(length(ind_suspicious_ch4)>0){
  message("the following rows show suspiciously HIGH methane levels")
  as.data.frame(fieldsheet[ind_suspicious_ch4,c("pilot_site","subsite","person_sampling","plot_id","initial_ch4","final_ch4")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_suspicious_ch4, "suspiciously low CH4")
}

```



## Quality check 4 - suspicious depth values

```{r Depth values, echo=F}
# missing depth values
ind_NA_depth <- which(is.na(fieldsheet$water_depth))
if(length(ind_NA_depth)>0){
  message("the following rows show NA water depths:")
  as.data.frame(fieldsheet[ind_NA_depth,c("pilot_site","subsite","person_sampling","plot_id","water_depth")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_NA_depth, "NAs in water depths")
}

# --------- rows with depth possibly reported in m instead of cm
fieldsheet_not_dry <- fieldsheet[fieldsheet$water_depth != 0,]
ind_suspicious_depth <- which(fieldsheet_not_dry$water_depth < 2)

if(length(ind_suspicious_depth)>0){
  message("the following rows show suspicious water depth ranges:")
  as.data.frame(fieldsheet_not_dry[ind_suspicious_depth,c("pilot_site","subsite","person_sampling","plot_id","water_depth")])
  fieldsheet_flagged <- append2table(fieldsheet_not_dry, fieldsheet_flagged, ind_suspicious_depth, "supsiciously low water depth")
}
if(length(ind_suspicious_depth)>0){
  message("These rows correspond to the following subsites:")
  unique(fieldsheet_not_dry$subsite[ind_suspicious_depth])
}
```



## Quality check 5 - missing logger data information

```{r Loggers, echo=F}
ind_loggers <- which(is.na(fieldsheet$logger_floating_chamber) | is.na(fieldsheet$logger_transparent_chamber))
if(length(ind_loggers)>0){
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_loggers, "No data logger info")
  message("the following subsites do not have logger data info")
  unique(fieldsheet$subsite[ind_loggers])
}
```




## Quality check 6 - Possible mistake with chamber type or chamber_height

```{r chambers, echo=F}

ind_suspicious_chamb_type <- which(fieldsheet$chamber_type=="tube"& fieldsheet$chamber_height_cm==0 & fieldsheet$strata=="open water")
if(length(ind_suspicious_chamb_type)>0){
  message(" Possible mistake with chamber type or chamber_height")
  as.data.frame(fieldsheet[ind_suspicious_chamb_type,c("pilot_site","subsite","person_sampling","plot_id","chamber_type","strata","chamber_height_cm")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_suspicious_chamb_type, "suspicious chamber type or chamber_height")
}


ind_suspicious_chamb_type <- which(fieldsheet$chamber_type=="floating" & fieldsheet$strata!="open water")
if(length(ind_suspicious_chamb_type)>0){
  message("floating chamber used outside open water")
  as.data.frame(fieldsheet[ind_suspicious_chamb_type,c("pilot_site","subsite","person_sampling","plot_id","chamber_type","strata","chamber_height_cm")])
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_suspicious_chamb_type, "floating chamber used outside open water")
}
```



## Quality check 7 - Possible mistake with coordinates

```{r coordinates, warning=FALSE, echo=F}

my_shp <- st_as_sf(fieldsheet,
                   coords = c("longitude", "latitude"),
                   crs = 4326)
# setwd(paste0(dropbox_root,"/GIS"))
# writing shp file to GIS directory
# myfilename <- paste0("GHG_sampling_points")
# st_write(my_shp, dsn = myfilename, layer = myfilename, driver = "ESRI Shapefile", append = F)

# load subsites polygons
setwd(dropbox_root)
shp_sites <- st_read("./GIS/Shape_files_subsites/Rough shapes for export/Subsites_rough_forExport.shp")

# intersect my_shp with shp_sites and find points outside polygons
# find corresponding basins
polygs = st_as_sf(shp_sites)
points = st_as_sf(my_shp)
sf_use_s2(FALSE)

intrsct = st_intersection(points, polygs) # intersect polygons with points, keeping the information from both

ind_match <- match(x = points$uniqID, table = intrsct$uniqID)
if(sum(is.na(ind_match))>0){
  message("Wrong coordinates")
  print(points$uniqID[is.na(ind_match)])
  
  ind_wrongXY <- match( points$uniqID[is.na(ind_match)], fieldsheet$uniqID)
  fieldsheet_flagged <- append2table(fieldsheet, fieldsheet_flagged, ind_wrongXY, "wrong XY coordinates")
}
```



## Recap table with all issues flagged

```{r Recap, warning=FALSE, echo=F}


setwd(paste0(dropbox_root,"/GHG/Working data/Fieldsheets_Quality"))
date_now <- Sys.Date()
time_now <- Sys.time()
myfilename <- paste("fieldsheets_quality_check",as.character(date_now),hour(time_now), minute(time_now),sep = "_")
write.csv(x = fieldsheet_flagged, file = paste0(myfilename,".csv"), row.names = F)



print(as.data.frame(fieldsheet_flagged[,c("uniqID","person_sampling","flag")]))
```

